---
class: inverse, center, middle

# Online Firestorms in Brand Communities 

---
class: font160
# Motivation 

**Strategic Question**: How can a firm detect, prevent and mitigate online firestorms in brand communities

* **Online Firestorm**: negative OWoM that gets substantial support from other customers quickly


**Why relevant?** 

* 72% of firms rate preparedness for online firestorms as below average
* Limited understanding at how to deal with negative OWoM 
* Difficulty predicting what happens when negative OWoM propagates through a large audience

---
# Online Firestorm Erupting 

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_eruption.png"
knitr::include_graphics(url)
```

---
# Online Firestorm Prevention 


```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_prevention.png"
knitr::include_graphics(url)
```

---
# Online Firestorm Mitigation


```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_mitigation.png"
knitr::include_graphics(url)
```

---
class: font140
# Data 

**Source**: Facebook Brand Communities for S&P 500 companies 

**Date Range**: October 2011 to Jan 2016

**Sample**: 472,995 negative posts across 89 online brand communities

**Important Definitions**:

* SST = strength of structural ties
  * Likes + comments $\times$ Given / received in community by a poster 

* LSM = linguistic similarity
  * How similar is complaining customers language with that of the existing posts in brand community

**Lurking question?** Firestorms vs Product Harm crises? 

* Firestorms emerge and end usually within 24 hours

---
# The Paper in a Nutshell

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_paper_nutshell.png"
knitr::include_graphics(url)
```

---
# Predicting Firestorms

```{r, echo = FALSE, fig.align = "center", out.width="80%"}
url <- "figs/of_model_prediction.png"
knitr::include_graphics(url)
```

---
# Prevention 

```{r, echo = FALSE, fig.align = "center", out.width="70%"}
url <- "figs/of_model_prevention.png"
knitr::include_graphics(url)
```

---
# Mitigation

```{r, echo = FALSE, fig.align = "center", out.width="60%"}
url <- "figs/of_model_mitigation.png"
knitr::include_graphics(url)
```