Detecting, Preventing, and Mitigating Online Firestorms in Brand Communities 

* [Herhausen, Ludwig, Grewal, Wulf & Schoegel (2019)](https://journals.sagepub.com/doi/10.1177/0022242918822300)

---
class: inverse, center, middle

# Online Firestorms in Brand Communities 

---
class: font160
# Motivation 

**Strategic Question**: How can a firm detect, prevent and mitigate online firestorms in brand communities

* **Online Firestorm**: negative OWoM that gets substantial support from other customers quickly


**Why relevant?** 

* 72% of firms rate preparedness for online firestorms as below average
* Limited understanding at how to deal with negative OWoM 
* Difficulty predicting what happens when negative OWoM propagates through a large audience

---
# Online Firestorm Erupting 

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_eruption.png"
knitr::include_graphics(url)
```

---
# Online Firestorm Prevention 


```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_prevention.png"
knitr::include_graphics(url)
```

---
# Online Firestorm Mitigation


```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_mitigation.png"
knitr::include_graphics(url)
```

---
class: font140
# Data 

**Source**: Facebook Brand Communities for S&P 500 companies 

**Date Range**: October 2011 to Jan 2016

**Sample**: 472,995 negative posts across 89 online brand communities

**Important Definitions**:

* SST = strength of structural ties
  * Likes + comments $\times$ Given / received in community by a poster 

* LSM = linguistic similarity
  * How similar is complaining customers language with that of the existing posts in brand community

**Lurking question?** Firestorms vs Product Harm crises? 

* Firestorms emerge and end usually within 24 hours

---
# The Paper in a Nutshell

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_paper_nutshell.png"
knitr::include_graphics(url)
```

---
# Predicting Firestorms

```{r, echo = FALSE, fig.align = "center", out.width="80%"}
url <- "figs/of_model_prediction.png"
knitr::include_graphics(url)
```

---
# Prevention 

```{r, echo = FALSE, fig.align = "center", out.width="70%"}
url <- "figs/of_model_prevention.png"
knitr::include_graphics(url)
```

---
# Mitigation

```{r, echo = FALSE, fig.align = "center", out.width="60%"}
url <- "figs/of_model_mitigation.png"
knitr::include_graphics(url)
```


---
class: font150
# Summary 

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/of_summary.png"
knitr::include_graphics(url)
```

Remark: Everything here is **correlation**

* Don't be swayed by authors want to interpret causally 

---
class: font160
# Motivation 

**Strategic Question**: How to design ads to *seed* a viral advertising campaign

Dimensions studied:

* Creativity: intruguing, entertaining, gains consumer's attention 
* Informativeness: provides needed information or essential facts

**Why relevant?** 

* Dimensions under a firm's control 
* At the expense of advertising effectiveness 


---
# Conceptual Framework 

<br>

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/propogate_conceptual.png"
knitr::include_graphics(url)
```

---
class: font150
# Flow of the Paper: Lab Studies 

Multiple studies $\rightarrow$ "Typical" consumer behaviour style paper

**Study 1** Field data: Views of YouTube ads

* Do certain types of ads generate more virality in "real" data?

**Study 2**: Real Ads *and* circulation intentions 

* For propagators vs non-propagators 

**Study 3**: Vary creativity and informativeness of a focal ad

* Control the ad content
* Remove omitted variable bias driven by product category of the real ads

---
class: font160
# Study 1: Results

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/propogate_results.png"
knitr::include_graphics(url)
```

* Creativity and informativeness positively correlated with views
* Propagators too 
  * Measured by number of comments ... so take this with a grain of salt


---
class: font140
# Study 2: Results

```{r, echo = FALSE, fig.align = "center", out.width="70%"}
url <- "figs/propogate_study2_results.png"
knitr::include_graphics(url)
```

* Higher creativity associated with higher intent to disseminate
* Propogators even more likely when highly informative and highly creative

*Remark*: Propagator defined as "having friends" and "talking to them a lot"

---
# Study 3: Manipulating the Ad Content

```{r, echo = FALSE, fig.align = "center", out.width="80%"}
url <- "figs/propogate_study3_stimuli.png"
knitr::include_graphics(url)
```

---
class: font140
# Study 3: Results

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/propogate_results.png"
knitr::include_graphics(url)
```

Dissemination: 

* High creativity positively associated with intent to share
* High creativity and highly informative even more so for propagators 

Social Status elevation: (a mechanism)

* Propagators do this to elevate their social status

---
class: font160
# Takeaways

* Ads more likely to be shared when they are creative and informative 
  * Even more likely for more socially connected individuals

* Driven by socially connected individuals desire to elevate (perceived) social status 
